---
title: "01-B-BHM-Simstudy-ORR-report"
author: "Sandro Gsteiger"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 7
    highlight: 'default'
    reference_docx: word_template.docx
    toc: yes
    toc_depth: 4
  # html_document:
  #   toc: true
  #   toc_float: true
  #   number_sections: true
  # editor_options: 
  #   chunk_output_type: console
knit: (function(inputFile, encoding) {rmarkdown::render(inputFile, encoding = encoding, output_dir = "outputs")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list = ls())

library(dplyr)
library(tidyr)
library(ggplot2)

for(s in dir("R")){
  source(file.path(".", "R", s))
}

# load outputs generated with the simulation script
sim_files <- c("01-B-global_par.RData",
               "01-B-res.RData",
               "01-B-scenario-grid.RData")
for(s in sim_files){
  load(file = paste("outputs/", s, sep = ""))  
}
```

# Introduction

## General description
Simulation study to assess the predictive uncertainty in a binary endpoint analyzed with a Bayesian hierarchical model as a function of number of groups and sample size per group.

How perform different design choices (number of groups, size of groups, larger lead group or not)?

What about the analysis model (HN(sigma=1), HN(sigma=0.5), or U(0,5) prior)?

## Defining and interpreting true levels of heterogeneity - how to define a meaningful (analysis) prior for the RE SD?

Assume the true population event rate is $p_0$ and the between-tissue-heterogeneity is $\tau$. Let $p_i$ be the true rate for tissue $i$ and $\theta_i = \textrm{logit}(p_i)$. The random effects model assumes exchangeability on the logit scale, this means
$$ \theta_i \sim N(\theta_0, \tau^2).$$

The range of true tissue specific event rates $p_i$ as well as threshold probabilities will depend on $\tau$.

__Figure__ 95% and 50% ranges of tissue specific event rates as a function of heterogeneity (random effects SD)
```{r}
# calculate the prediction intervals as a function of the re.sd
p.pop <- 0.5
re.sd <- matrix(seq(0, 5, 0.1), nrow = 1)
q <- matrix(qnorm(c(0.025, 0.25, 0.75, 0.975), 0, 1), ncol = 1)
theta_mat <- q %*% re.sd
p_mat <- boot::inv.logit(theta_mat)
row.names(p_mat) <- c("PrIlo", "Q1", "Q3", "PrIup")
dhet <- data.frame(re.sd = t(re.sd),
                   t(p_mat)
                   )
dhet_l <- gather(dhet, 
                 key = "key", 
                 value = "value", "PrIlo", "Q1", "Q3", "PrIup", -re.sd)

diqr <- dhet %>% 
  filter(re.sd <= 1) %>%
  select(re.sd, Q1, Q3)

dpri <- dhet %>% 
  filter(re.sd <= 1) %>%
  select(re.sd, PrIlo, PrIup)


# make the plot
fig <- ggplot() +
  geom_line(data = dhet_l %>% filter(key %in% c("Q1", "Q3")), aes(re.sd, value, group = key), linetype = "dashed") +
  geom_line(data = dhet_l %>% filter(key %in% c("PrIlo", "PrIup")), aes(re.sd, value, group = key), linetype = "dotted") +
  geom_linerange(data = dpri, aes(x = re.sd, ymin = PrIlo, ymax = PrIup), size = 1) +
  geom_linerange(data = diqr, aes(x = re.sd, ymin = Q1, ymax = Q3), size = 2) +
  theme_bw() +
  scale_x_continuous(limits = c(0, 1.01), breaks = seq(0, 1, 0.1), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1), expand = c(0, 0)) +
  ylab("Event probability") +
  xlab("Random effects SD")
fig

rm(fig)
```

__Figure__ idem (alternative version)
```{r}
fig <- ggplot(data = dhet, aes(x = re.sd)) +
  geom_ribbon(aes(ymin = PrIlo, ymax = PrIup), fill = "grey60", alpha = 0.5) +
  geom_ribbon(aes(ymin = Q1, ymax = Q3), fill = "grey40") +
  geom_linerange(aes(ymin = PrIlo, ymax = PrIup), size = 1) +
  geom_linerange(aes(ymin = Q1, ymax = Q3), size = 2) +
  theme_bw() +
  scale_x_continuous(limits = c(0, 1.01), breaks = seq(0, 1, 0.1), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1), expand = c(0, 0)) +
  ylab("Event probability") +
  xlab("Random effects SD")
fig

rm(fig)
```



__Table__ 95% and 50% ranges of tissue specific event rates for selected levels of heterogeneity (random effects SD) as well as probability of being below 0.3
```{r, results='asis'}
vnear <- function(x, y){
  all <- outer(x, y, FUN = near)
  out <- apply(all, MAR = 1, FUN = any)
  return(out)
}
thres <- global_par$p.threshold
taus <- unique(res$re.sd)
lab <- paste("P(p.new<", thres, ")", sep = "")
dsum <- round(data.frame(
  dhet[vnear(dhet$re.sd, taus),],
  pnorm((boot::logit(thres) - boot::logit(p.pop)) / taus, mean = 0, sd = 1), check.names = FALSE), 2)
colnames(dsum)[ncol(dsum)] <- lab
  
pander::pandoc.table(dsum, row.names = FALSE, split.tables = Inf, justify = "left")
```



__Figure__ 50% prediction intervals (dark grey) and 95% prediction intervals (light grey) for true tissue specific event probabilities as a function of RE SD
```{r}
fig <- ggplot(data = dhet, aes(x = re.sd)) +
  geom_ribbon(aes(ymin = PrIlo, ymax = PrIup), fill = "grey60", alpha = 0.5) +
  geom_ribbon(aes(ymin = Q1, ymax = Q3), fill = "grey40") +
  geom_linerange(data = dhet %>% filter(re.sd %in% seq(0.5, 4.5, 0.5)), 
                 aes(ymin = PrIlo, ymax = PrIup), size = 1) +
  geom_linerange(data = dhet %>% filter(re.sd %in% seq(0.5, 4.5, 0.5)),
                 aes(ymin = Q1, ymax = Q3), size = 2) +
  scale_x_continuous(limits = c(0, 5.05), breaks = seq(0, 5, 0.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2), expand = c(0, 0)) +
  ylab("Event probability") +
  xlab("Random effects SD") +
  theme_bw()
fig
```


__Figure__ 50% prediction intervals (dark grey) and 95% prediction intervals (light grey) for true tissue specific event probabilities as a function of RE SD along wtih different prior distributions
```{r}
# -> reuse fig from above and combine w figure below

# uniform prior
u_0_5  <- function(x){0.2}
hn_1   <- function(x){2 * dnorm(x, 0, 1)}
hn_0.5 <- function(x){2 * dnorm(x, 0, 0.5)}

dprior <- data.frame(x = seq(0, 5, 0.1)) %>%
  mutate("U(0,5)" = u_0_5(x),
         "HN(0.5)"= hn_0.5(x),
         "HN(1)"   = hn_1(x)) %>%
  gather(Prior, Density, 2:4)

fig1 <- ggplot(data = dprior, aes(x = x, y = Density)) +
  geom_area(aes(fill = Prior, linetype = Prior), 
            colour = 1, alpha = 0.5, position = "identity") +
  scale_x_continuous(limits = c(0, 5.05), breaks = seq(0, 5, 0.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 2), breaks = seq(0, 2, 0.5), expand = c(0, 0)) +
  xlab("Random effects SD") +
  theme_bw() +
  theme(legend.position = "bottom")
fig1

egg::ggarrange(fig, fig1, ncol = 1)

rm(fig, fig1)
```

# Simulation settings 

## Simulation scenarios
Simulations were run over the full grid of combinations below.
```{r, echo=TRUE}
# True rate and heterogeneity
bin_grid 

# Sample sizes of lead group and subsequent groups
size_grid

# Number of groups
grp_grid

# Analysis priors (RE SD)
prior_grid

# Resulting full scenario grid
dim(full_grid)
head(full_grid)
tail(full_grid)
```

## Global parameters
```{r, echo=TRUE}
global_par
```


# Simulation results

```{r}
# combine scenario characteristics with sim results for analysis
res <- res %>%
  left_join(data.frame(scenario = 1:nrow(full_grid),
                       full_grid),
            by = "scenario") %>%
  mutate(n.tot = lead.grp.size + (n.grp - 1) * subseq.grp.size) %>%
  mutate(n_first = lead.grp.size, n_others = subseq.grp.size) # nicer labels

dim(res)
head(res)
tail(res)
```


## Estimating between-study-heterogeneity (RE SD)

### Posterior median RE SD ($\tau$)

__Figure__ Posterior median of $\tau$ (mean over `n global_par$n_sim` simulations)
```{r}
dat <- res %>%
  filter(node == "re.sd", var1 == "50%") %>%
  filter(var2 == "mean") 

for(s in unique(dat$re.sd)){
  fig <- ggplot(data = dat %>% filter(re.sd == s)) +
    geom_hline(aes(yintercept = s), linetype = "dashed") +
    geom_line(aes(x = n.grp, y = value, color = dist)) +
    facet_grid(n_first ~ n_others, labeller = label_both) +
    ggtitle(bquote(tau==.(s))) +
    xlab("Number of groups") + 
    ylab(expression(Mean~tau~estimate)) +
    theme_bw() +
    theme(legend.title = element_blank(), legend.position = "top") 
  print(fig)
  rm(fig)
}
```


**Mean Squared Error (MSE)** of the posterior median as an estimator for $\tau$, 
$$MSE = Var + bias^2.$$
__Figure__ MSE of the posterior median RE SD
```{r}
dat <- res %>%
  filter(node == "re.sd", var1 == "50%") %>%
  filter(var2 %in% c("mean", "sd")) %>%
  spread(var2, value) %>%
  mutate(MSE = sd^2 + (mean - re.sd)^2)

for(s in unique(dat$re.sd)){
  fig <- ggplot(data = dat %>% filter(re.sd == s)) +
    geom_line(aes(x = n.grp, y = MSE, color = dist)) +
    facet_grid(n_first ~ n_others, labeller = label_both) +
    ggtitle(bquote(tau==.(s))) +
    xlab("Number of groups") + 
    theme_bw() +
    theme(legend.title = element_blank(), legend.position = "top")
  print(fig)
  rm(fig)
}
```


## Predictive uncertainty: new tissue response rate


--- HERE ---

__Figure__ Length of 95% prediction interval of new group event probability (mean over `n global_par$n_sim` simulations)
```{r}

```


__Figure__ Inter-quartile range of prediction interval of new group event probability (mean over `n global_par$n_sim` simulations)
```{r}

```

__Figure__ Estimated posterior probability of new group event probability being lower than `r global_par$p.threshold`, P(p.new<`r global_par$p.threshold`)  (mean over `n global_par$n_sim` simulations)
```{r}

```








# Session info
```{r}
getwd()
sessionInfo()
```
