---
title: "01-B3-BHM-Simstudy-ORR-report"
author: "Sandro Gsteiger"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 7
    highlight: 'default'
    reference_docx: word_template.docx
    toc: yes
    toc_depth: 4
  # html_document:
  #   toc: true
  #   toc_float: true
  #   number_sections: true
  # editor_options: 
  #   chunk_output_type: console
knit: (function(inputFile, encoding) {rmarkdown::render(inputFile, encoding = encoding, output_dir = "outputs")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list = ls())

library(dplyr)
library(tidyr)
library(ggplot2)

for(s in dir("R")){
  source(file.path(".", "R", s))
}

# load outputs generated with the simulation script
sim_files <- c("01-B3-global_par.RData",
               "01-B3-res.RData",
               "01-B3-scenario-grid.RData")
for(s in sim_files){
  load(file = paste("outputs/", s, sep = ""))  
}

# combine scenario characteristics with sim results for analysis
res <- res %>%
  left_join(data.frame(scenario = 1:nrow(full_grid),
                       full_grid),
            by = "scenario") %>%
  mutate(n.tot = lead.grp.size + (n.grp - 1) * subseq.grp.size) %>%
  mutate(n_first = lead.grp.size, n_others = subseq.grp.size) # nicer labels
```

# Introduction


```{r}
# calculate the prediction intervals as a function of the re.sd
p.pop <- 0.5
re.sd <- matrix(seq(0, 5, 0.1), nrow = 1)
q <- matrix(qnorm(c(0.025, 0.25, 0.75, 0.975), 0, 1), ncol = 1)
theta_mat <- q %*% re.sd
p_mat <- boot::inv.logit(theta_mat)
row.names(p_mat) <- c("PrIlo", "Q1", "Q3", "PrIup")
dhet <- data.frame(re.sd = t(re.sd),
                   t(p_mat)
                   )
dhet_l <- gather(dhet, 
                 key = "key", 
                 value = "value", "PrIlo", "Q1", "Q3", "PrIup", -re.sd)

diqr <- dhet %>% 
  filter(re.sd <= 1) %>%
  select(re.sd, Q1, Q3)

dpri <- dhet %>% 
  filter(re.sd <= 1) %>%
  select(re.sd, PrIlo, PrIup)
```



__Table__ 95% and 50% ranges of tissue specific event rates for selected levels of heterogeneity (random effects SD) as well as probability of being below 0.3
```{r, results='asis'}
vnear <- function(x, y){
  all <- outer(x, y, FUN = near)
  out <- apply(all, MAR = 1, FUN = any)
  return(out)
}
thres <- global_par$p.threshold
taus <- unique(res$re.sd)
lab <- paste("ProbBelow", thres, sep = "")
dsum <- round(data.frame(
  dhet[vnear(dhet$re.sd, taus),],
  pnorm((boot::logit(thres) - boot::logit(p.pop)) / taus, mean = 0, sd = 1), check.names = FALSE), 2)
colnames(dsum)[ncol(dsum)] <- lab
  
pander::pandoc.table(dsum, row.names = FALSE, split.tables = Inf, justify = "left")
```



# Simulation settings 

## Simulation scenarios
Simulations were run over the full grid of combinations below.
```{r, echo=TRUE}
# True rate and heterogeneity
bin_grid 

# Sample sizes (lead group and subsequent groups) and group numbers
sample_grid


# Analysis priors (RE SD)
prior_grid

# Resulting full scenario grid
dim(full_grid)
head(full_grid)
tail(full_grid)
```


## Global parameters
```{r, echo=TRUE}
global_par
```


# Simulation results

Dimension, first and last few lines of results data frame.
```{r}
dim(res)
head(res)
tail(res)
```


## Estimating between-study-heterogeneity (RE SD)

### Posterior median RE SD ($\tau$)

__Figure__ Posterior median of $\tau$ (mean over `r global_par$n_sim` simulations)
```{r, fig.height=7}
dat <- res %>%
  filter(node == "re.sd", var1 == "50%") %>%
  filter(var2 == "mean") 

for(s in unique(dat$re.sd)){
  fig <- ggplot(data = dat %>% filter(re.sd == s)) +
    geom_hline(aes(yintercept = s), linetype = "dashed") +
    geom_line(aes(x = n.grp, y = value, color = dist)) +
    facet_grid(n_others ~ n_first, labeller = label_both) +
    ggtitle(bquote(tau==.(s))) +
    xlab("Number of groups") + 
    ylab(expression(Mean~tau~estimate)) +
    theme_bw() +
    theme(legend.title = element_blank(), legend.position = "top") 
  print(fig)
  rm(fig)
}
rm(dat)
```


### Mean Squared Error

**Mean Squared Error (MSE)** of the posterior median as an estimator for $\tau$, 
$$MSE = Var + bias^2.$$
__Figure__ MSE of the posterior median RE SD
```{r, fig.height=7}
dat <- res %>%
  filter(node == "re.sd", var1 == "50%") %>%
  filter(var2 %in% c("mean", "sd")) %>%
  spread(var2, value) %>%
  mutate(MSE = sd^2 + (mean - re.sd)^2)

for(s in unique(dat$re.sd)){
  fig <- ggplot(data = dat %>% filter(re.sd == s)) +
    geom_line(aes(x = n.grp, y = MSE, color = dist)) +
    facet_grid(n_others ~ n_first, labeller = label_both) +
    ggtitle(bquote(tau==.(s))) +
    xlab("Number of groups") + 
    theme_bw() +
    theme(legend.title = element_blank(), legend.position = "top")
  print(fig)
  rm(fig)
}
rm(dat)
```


## Predictive uncertainty: new tissue response rate

### 95% Prediction interval

```{r}
# calculate true range of 95% CrI and IQR
dsum <- dsum %>%
  mutate(IQR = Q3 - Q1, PrIR = PrIup - PrIlo)
```


__Figure__ Length of 95% prediction interval of new group event probability (mean over `r global_par$n_sim` simulations)
```{r, fig.height=7}
dat <- res %>%
  filter(node == "p.new", var1 == "CrIR") %>%
  filter(var2 %in% c("mean"))

for(s in unique(dat$re.sd)){
  fig <- ggplot(data = dat %>% filter(re.sd == s)) +
    geom_hline(data = dsum %>% filter(re.sd == s), aes(yintercept = PrIR), linetype = "dashed") +
    geom_line(aes(x = n.grp, y = value, color = dist)) +
    facet_grid(n_others ~ n_first, labeller = label_both) +
    ylim(0, 1) +
    ggtitle(bquote(tau==.(s))) +
    xlab("Number of groups") + 
    ylab("Length of 95% PrI") +
    theme_bw() +
    theme(legend.title = element_blank(), legend.position = "top")
  print(fig)
  rm(fig)
}

rm(dat)
```


### Interquartile range

__Figure__ Inter-quartile range of prediction interval of new group event probability (mean over `r global_par$n_sim` simulations)
```{r, fig.height=7}
dat <- res %>%
  filter(node == "p.new", var1 == "IQR") %>%
  filter(var2 %in% c("mean"))

for(s in unique(dat$re.sd)){
  fig <- ggplot(data = dat %>% filter(re.sd == s)) +
    geom_hline(data = dsum %>% filter(re.sd == s), aes(yintercept = IQR), linetype = "dashed") +
    geom_line(aes(x = n.grp, y = value, color = dist)) +
    facet_grid(n_others ~ n_first, labeller = label_both) +
    ylim(0, 0.6) +
    ggtitle(bquote(tau==.(s))) +
    xlab("Number of groups") + 
    ylab("IQR") +
    theme_bw() +
    theme(legend.title = element_blank(), legend.position = "top")
  print(fig)
  rm(fig)
}

rm(dat)
```


### Threshold probabilities

__Figure__ Estimated posterior probability of new group event probability being lower than `r global_par$p.threshold`, P(p.new<`r global_par$p.threshold`)  (mean over `r global_par$n_sim` simulations)
```{r, fig.height=7}
dat <- res %>%
  filter(node == "p.new", var1 == "ProbAbove0.3") %>%
  filter(var2 %in% c("mean")) %>%
  mutate(value = 1 - value)

for(s in unique(dat$re.sd)){
  fig <- ggplot(data = dat %>% filter(re.sd == s)) +
    geom_hline(data = dsum %>% filter(re.sd == s), aes(yintercept = ProbBelow0.3), linetype = "dashed") +
    geom_line(aes(x = n.grp, y = value, color = dist)) +
    facet_grid(n_others ~ n_first, labeller = label_both) +
    ylim(0, 0.4) +
    ggtitle(bquote(tau==.(s))) +
    xlab("Number of groups") + 
    ylab("P(p.new<0.3)") +
    theme_bw() +
    theme(legend.title = element_blank(), legend.position = "top")
  print(fig)
  rm(fig)
}

```

#### Compare impact of group size vs number of groups

__Figure__ Threshold probability as a function of total sample size. The panels give the subsequent group size (the first group has always a size of 5).
```{r, fig.height = 4}
dat <- res %>%
  filter(node == "p.new", var1 == "ProbAbove0.3") %>%
  filter(var2 %in% c("mean")) %>%
  mutate(value = 1 - value)

for(s in unique(dat$re.sd)){
  for(p in unique(dat$dist)){
    fig <- ggplot(data = dat %>% filter(re.sd == s, dist == p)) +
      geom_hline(data = dsum %>% filter(re.sd == s), 
                 aes(yintercept = ProbBelow0.3), linetype = "dashed") +
      geom_line(aes(x = n.tot, y = value)) +
      facet_grid(.~n_others) +
      ylim(0, 0.4) +
      ggtitle(bquote(tau==.(s)), subtitle = p) +
      xlab("Total sample size") + 
      ylab("P(p.new<0.3)") +
      theme_bw() +
      theme(legend.position = "top")
    print(fig)
    rm(fig)
  }
}

```



# Session info
```{r}
getwd()
sessionInfo()
```
