---
title: "02-B-VoI-Casestudy-pop-vs-pred"
author: "Sandro Gsteiger"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  word_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 7
    highlight: 'default'
    reference_docx: word_template.docx
    toc: yes
    toc_depth: 4
  # html_document:
  #   toc: true
  #   toc_float: true
  #   number_sections: true
  # editor_options: 
  #   chunk_output_type: console
knit: (function(inputFile, encoding) {rmarkdown::render(inputFile, encoding = encoding, output_dir = "outputs")})
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE)
options(tibble.width = Inf)

library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(BCEA)

for(s in dir("R")){
  source(file.path(".", "R", s))
}

r_seed <- 3947623
set.seed(r_seed)
```

# Description
Use very simple, stylized desision model to assess the relevance of predictive uncertainty with tumour agnostic therapies.

Compare EVPI (and possibly EVPPI) for two decision rules:

1. Use what might represent single-tissue (or population?) estimates for decision making (and their uncertainty.
2. Use what might represent the predicted outcomes in a new tissue in the decision rule.


# Stylized decision model

Partitioned survival model.

Weibull model for OS and PFS, shape parameter assumed the same. Shape parameter estimated from Larotrectinib OS data. Scale paramters (with uncertainty) estimated from Larotrectinib 12-months survival data (OS and PFS).

Assessment of impact of predictive uncertainty (new tissue type).

Implementation by adapting functions publically available from DARTH group
(https://github.com/DARTH-git/Partitioned-Survival-Analysis).


# Base case analysis

## Model inputs

* Baseline survival rates (new treatment): OS and PFS rates and their standard errors roughly taken from Laro data 
* Hazard ratios to obtain comparator group: assumptions (set to obtain "workable" model for case study, i.e. new intervention reasonnable likelihood of being cost effective)
* Utilities: assumptions based on from lit
  * Pre-pog: smal diff. between treatments
  * Post-progression: 
    + no difference between the two groups
    + this was the preferred assumption by the NICE committee in the Laro dossier
    + the manufacturer and some clinical experts said difference in post-prog. utility were plausible
* Costs:
  * New trt, pre-prog: 
    * per pERC report about 20k CAN $ per month; -> 12k GBP
    * per NICE ACM2: list price GBP 15k per month (then confidential discount)
  * Ref trt, pre-prog: per pERC about 2k CAN $, though extreme between tissue variability
  * Post-progression state (indep. of trt); assumption
  * Make rough assumption for total costs per cycle (reflecting also drug administration, other direct and indirect medical costs such as for ecample those induced by AEs etc.)
* Discount rate: 3.5% for costs and outcomes per NHS/NICE recommendation
* Time-horizon: 30 years.


Note that _time scale_ is _years_ (but _cycle length_ is _months_).

```{r, echo=TRUE}
model_pars <- list(
  shape       = 1.2,
  OS_scale  = 0.128,      
  PFS_scale = 0.446,
  HR_OS       = 0.3,
  HR_PFS      = 0.3,
  u_pre_trt   = 0.8,      # utility in pre-progrsssion state, new treatment
  u_pre_ref   = 0.7,      # idem, reference treatment
  u_post      = 0.6,     # utility, post-progression state
  u_death     = 0,
  c_pre_trt   =  4500,    # costs in pre-progrsssion state per cycle, new treatment
  c_pre_ref   =  3000,    # idem, reference treatment
  c_post      =  1000,    # costs per cycle, post-progression state
  c_death     =     0     
)

model_settings <- list(
  state_names    = c("Pre-progression", "Post-progression", "Death"),
  horizon        = 30,
  cycle_length   = 1 / 12,  # cycle length 1-month (time scale: years) 
  disc_rate      = 0.035    # rate per year
)
```


## Visualize the resulting survival models
```{r}
pars_trt <- list(
  scale.os  = model_pars$OS_scale,
  scale.pfs = model_pars$PFS_scale,
  shape     = model_pars$shape
)

pars_ref <- list(
  scale.os  = model_pars$OS_scale * (1 / model_pars$HR_OS),
  scale.pfs = model_pars$PFS_scale * (1 / model_pars$HR_PFS),
  shape     = model_pars$shape
)


t_eval <- seq(0, model_settings$horizon, by = model_settings$cycle_length) # time scale: years
```


```{r}
dtte_trt <- data.frame(years = t_eval,
                       intervention = "New therapy") %>%
  mutate(OS  = exp(-pars_trt$scale.os  * years ^ pars_trt$shape),
         PFS = exp(-pars_trt$scale.pfs  *years ^ pars_trt$shape))

dtte_ref <- data.frame(years = t_eval,
                       intervention = "Reference") %>%
  mutate(OS  = exp(-pars_ref$scale.os  * years ^ pars_ref$shape),
         PFS = exp(-pars_ref$scale.pfs  *years ^ pars_ref$shape))

dtte <- rbind(dtte_trt, dtte_ref)

```


__Figure__ Resulting OS Weibull models
```{r, message=FALSE, warning=FALSE}
ggplot(data = dtte) +
  geom_line(aes(years, OS, color = intervention, linetype = intervention), size = 1.2) +
  scale_x_continuous(breaks = 0:20, limits = c(0, 20), expand = c(0, 0)) +
  scale_y_continuous(breaks = 0:10/10, expand = c(0, 0)) +
  theme_bw() +
  theme(legend.title = element_blank(), legend.position = "top")
```


__Figure__ Resulting PFS Weibull models
```{r}
ggplot(data = dtte) +
  geom_line(aes(years, PFS, color = intervention, linetype = intervention), size = 1.2) +
  scale_x_continuous(breaks = 0:20, limits = c(0, 20), expand = c(0, 0)) +
  scale_y_continuous(breaks = 0:10/10, expand = c(0, 0)) +
  theme_bw() +
  theme(legend.title = element_blank(), legend.position = "top")
```


## Economic evaluation

**Figure** Markov traces (base case analysis)
```{r}
trace_trt <- get_partsurv(t_eval, pars_trt)
trace_ref <- get_partsurv(t_eval, pars_ref)

trace_trt2 <- trace_trt %>%
  mutate(time = t_eval) %>%
  gather("State", "Proportion", -time)

trace_ref2 <- trace_ref %>%
  mutate(time = t_eval) %>%
  gather("State", "Proportion", -time)

ggplot(data = trace_trt2) +
  geom_line(aes(time, Proportion, color = State, linetype = State)) +
  xlab("Years") +
  ggtitle("New treatment") +
  xlim(0, 20) +
  theme_bw()

ggplot(data = trace_ref2) +
  geom_line(aes(time, Proportion, color = State, linetype = State)) +
  xlab("Years") +
  ggtitle("Reference treatment") +
  xlim(0, 20) +
  theme_bw()
```



Calculate total discounted costs and QALYs for the new and the reference treatments.
```{r}
ee_trt <- get_ee(t_eval, 
                 trace_trt,
                 costs_per_cycle = c(model_pars$c_pre_trt, model_pars$c_post, model_pars$c_death),
                 utils = c(model_pars$u_pre_trt, model_pars$u_post, model_pars$u_death),
                 dr = model_settings$disc_rate,
                 cycle_length = model_settings$cycle_length
                 )

ee_ref <- get_ee(t_eval, 
                 trace_ref,
                 costs_per_cycle = c(model_pars$c_pre_ref, model_pars$c_post, model_pars$c_death),
                 utils = c(model_pars$u_pre_ref, model_pars$u_post, model_pars$u_death),
                 dr = model_settings$disc_rate,
                 cycle_length = model_settings$cycle_length
                 )

```

** Model outputs (base case)**
```{r, echo=TRUE}
## new treatment
ee_trt

## ref treatment
ee_ref


## incremental costs and QALYs
delta <- ee_trt - ee_ref
names(delta) <- paste("delta.", names(delta), sep = "")
delta

## ICER
ICER <- c(costs.per.QALY = delta[[1]] / delta[[2]])
ICER
```


# PSA: population model

## Distributions used for parameter uncertainty

```{r, echo=TRUE}
# OS_scale parameter, new trt:
OS_scale_lsd <- (log(0.211) - log(0.0513)) / 3.92
OS_scale_lsd
dscale <- data.frame(x = seq(0.01, 0.3, 0.01)) %>%
  mutate(y = dlnorm(x, meanlog = log(model_pars$OS_scale), sdlog = OS_scale_lsd))
qplot(x, y, data = dscale, geom = "line") + 
  labs(title = "OS", x = "scale", y = "")

qlnorm(meanlog = log(model_pars$OS_scale), sdlog = OS_scale_lsd, p = c(0.025, 0.975))
```


```{r, echo=TRUE}
# PFS_scale parameter, new trt:
qnorm(mean = 0.446, sd = (0.635 - 0.288)/6, p = c(0.025, 0.975))

PFS_scale_lsd <- (log(0.635) - log(0.288)) / 3.92
PFS_scale_lsd
dscale <- data.frame(x = seq(0.2, 0.7, 0.01)) %>%
  mutate(y = dlnorm(x, meanlog = log(model_pars$PFS_scale), sdlog = PFS_scale_lsd))
qplot(x, y, data = dscale, geom = "line") + 
  labs(title = "PFS", x = "scale", y = "")

qlnorm(meanlog = log(model_pars$PFS_scale), sdlog = PFS_scale_lsd, p = c(0.025, 0.975))
```


```{r, echo=TRUE}
# HR OS, PFS ~ Beta(30, 70) => Expectation: 3/10
qbeta(30, 70, p = c(0.025, 0.975))
```


```{r, echo=TRUE}
# costs pre-progression state new treatment
# median: 
model_pars$c_pre_trt
qlnorm(meanlog = log(model_pars$c_pre_trt), sdlog = 0.05, p = c(0.025, 0.975))
```


```{r, echo=TRUE}
# costs pre-progression state ref treatment
# median: 
model_pars$c_pre_ref
qlnorm(meanlog = log(model_pars$c_pre_ref), sdlog = 0.05, p = c(0.025, 0.975))
```


## Run the PSA and plot the cost-effectiveness plane

```{r}
n_sim <- 1000

model_psa <- matrix(NA, ncol = (length(model_pars) + 7), nrow = n_sim)
colnames(model_psa) <- c(names(model_pars), 
                         "costs.trt", "QALYs.trt", "costs.ref", "QALYs.ref", 
                         "dCost", "dQALY", "ICER")
for(i in 1:n_sim){
  model_psa[i, "shape"]    <- model_pars$shape
  model_psa[i, "OS_scale"] <- rlnorm(1, meanlog = log(model_pars$OS_scale), sdlog = OS_scale_lsd)
  model_psa[i, "PFS_scale"]<- rlnorm(1, meanlog = log(model_pars$PFS_scale), sdlog = PFS_scale_lsd)
  model_psa[i, "HR_OS"]    <- rbeta(1, shape1 = 30, shape2 = 70)
  model_psa[i, "HR_PFS"]   <- rbeta(1, shape1 = 30, shape2 = 70)
  model_psa[i, "u_pre_trt"]<- runif(1, model_pars$u_pre_trt - 0.05, model_pars$u_pre_trt + 0.05)
  model_psa[i, "u_pre_ref"]<- runif(1, model_pars$u_pre_ref - 0.05, model_pars$u_pre_ref + 0.05)
  model_psa[i, "u_post"]   <- runif(1, model_pars$u_post - 0.05, model_pars$u_post + 0.05)
  model_psa[i, "u_death"]  <- model_pars$u_death
  model_psa[i, "c_pre_trt"]<- rlnorm(1, meanlog = log(model_pars$c_pre_trt), sdlog = 0.05)
  model_psa[i, "c_pre_ref"]<- rlnorm(1, meanlog = log(model_pars$c_pre_ref), sdlog = 0.05)
  model_psa[i, "c_post"]   <- model_pars$c_post
  model_psa[i, "c_death"]  <- model_pars$c_death

  # Calculate the Markov traces
  pars_trt_i <- list(
    scale.os  = model_psa[i, "OS_scale"],
    scale.pfs = model_psa[i, "PFS_scale"],
    shape     = model_psa[i, "shape"]
  )
  trace_trt_i <- get_partsurv(t_eval, pars_trt_i)

  pars_ref_i <- list(
    scale.os  = model_psa[i, "OS_scale"] * (1 / model_psa[i, "HR_OS"]),
    scale.pfs = model_psa[i, "PFS_scale"]* (1 / model_psa[i, "HR_PFS"]),
    shape     = model_psa[i, "shape"]
  )
  trace_ref_i <- get_partsurv(t_eval, pars_ref_i)

  # Do the economic evaluation
  ee_trt_i <- get_ee(t_eval, 
                   trace_trt_i,
                   costs_per_cycle = model_psa[i, c("c_pre_trt", "c_post", "c_death")],
                   utils           = model_psa[i, c("u_pre_trt", "u_post", "u_death")],
                   dr              = model_settings$disc_rate,
                   cycle_length    = model_settings$cycle_length
  )
  
  ee_ref_i <- get_ee(t_eval, 
                     trace_ref_i,
                     costs_per_cycle = model_psa[i, c("c_pre_ref", "c_post", "c_death")],
                     utils           = model_psa[i, c("u_pre_ref", "u_post", "u_death")],
                     dr              = model_settings$disc_rate,
                     cycle_length    = model_settings$cycle_length
  )
  
  # Store results
  model_psa[i, c("costs.trt", "QALYs.trt")] <- ee_trt_i
  model_psa[i, c("costs.ref", "QALYs.ref")] <- ee_ref_i
  model_psa[i, c("dCost", "dQALY")]         <- ee_trt_i - ee_ref_i
  model_psa[i, c("ICER")]                   <- model_psa[[i, "dCost"]] / model_psa[[i, "dQALY"]]
}

ggplot(data = as.data.frame(model_psa)) +
  geom_point(aes(dQALY, dCost), shape = 1) +
  geom_point(data = as.data.frame(t(delta)), 
             aes(delta.total.QALYs, delta.total.costs), col = "red") +
  geom_hline(yintercept = 0, lwd = 1.5) +
  geom_vline(xintercept = 0, lwd = 1.5) +
  geom_abline(intercept = 0, slope = 50000, linetype = "dashed") +
  scale_x_continuous(limits = c(-1, 5.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(-50000, 300000), 
                     breaks = seq(-50000, 300000, 50000), expand = c(0, 0)) +
  theme_bw()
```


## Summarize PSA outputs and (re)analyze the PSA outputs with the BCEA package
```{r, echo=TRUE}
# PSA outputs
e <- model_psa[, c("QALYs.trt", "QALYs.ref")] # ref in bcea terminology is our trt!!
c <- model_psa[, c("costs.trt", "costs.ref")] # caution to this !!
me <- apply(e, MAR = 2, FUN = mean)
me
mc <- apply(c, MAR = 2, FUN = mean)
mc

diff(mc) / diff(me)

# Compare with basecase analysis
ee_trt
ee_ref
delta
ICER


# re-analyze with bcea
res <- bcea(e, c, interventions = c("ref", "trt"), 
            Kmax = 100000, wtp = seq(0, 100000, 2000))

summary.bcea(res, wtp = 50000)
```

```{r}
plot(res, wtp = 50000)
```

## Calculate EVPPI
```{r}
par(mfrow = c(1,1))
# tte pars
voi <- evppi(c("HR_OS", "HR_PFS"), model_psa, res)
plot(voi)

voi <- evppi(c("OS_scale", "PFS_scale"), model_psa, res)
plot(voi)


# costs
voi <- evppi(c("c_pre_trt", "c_pre_ref"), model_psa, res)
plot(voi)

# utils
voi <- evppi(c("u_pre_trt", "u_pre_ref"), model_psa, res)
plot(voi)
```


# Redo PSA, but for predictive distribution (new tissue) - add uncertainty

Add predictive uncertainty wrt HRs, and costs.

Keep the baseline model (which per se would be expected to vary too!).

## Run the PSA and plot the cost-effectiveness plane
Summaries of distributions used for parameter uncertainty. In particular with respect to reference treatment.
```{r, echo=TRUE}
# HR OS, PFS ~ Beta(1.5, 3.5) => Expectation: 3/10, but very uncertain
qbeta(1.5, 3.5, p = c(0.025, 0.975))

# costs pre-progression state new treatment
qlnorm(meanlog = log(model_pars$c_pre_trt), sdlog = 0.3, p = c(0.025, 0.25, 0.5, 0.75, 0.975))

# costs pre-progression state ref treatment
# median: 1000
qlnorm(meanlog = log(model_pars$c_pre_ref), sdlog = 0.75, p = c(0.025, 0.25, 0.5, 0.75, 0.975))
```


```{r}
set.seed(r_seed)
n_sim <- 1000

model_psa2 <- matrix(NA, ncol = (length(model_pars) + 7), nrow = n_sim)
colnames(model_psa2) <- c(names(model_pars), 
                         "costs.trt", "QALYs.trt", "costs.ref", "QALYs.ref", 
                         "dCost", "dQALY", "ICER")
for(i in 1:n_sim){
  model_psa2[i, "shape"]    <- model_pars$shape
  model_psa2[i, "OS_scale"] <- rlnorm(1, meanlog = log(model_pars$OS_scale),  sdlog = OS_scale_lsd)
  model_psa2[i, "PFS_scale"]<- rlnorm(1, meanlog = log(model_pars$PFS_scale), sdlog = PFS_scale_lsd)
  model_psa2[i, "HR_OS"]    <- rbeta(1, shape1 = 1.5, shape2 = 3.5)
  model_psa2[i, "HR_PFS"]   <- rbeta(1, shape1 = 1.5, shape2 = 3.5)
  model_psa2[i, "u_pre_trt"]<- runif(1, model_pars$u_pre_trt - 0.05, model_pars$u_pre_trt + 0.05)
  model_psa2[i, "u_pre_ref"]<- runif(1, model_pars$u_pre_ref - 0.05, model_pars$u_pre_ref + 0.05)
  model_psa2[i, "u_post"]   <- runif(1, model_pars$u_post - 0.05, model_pars$u_post + 0.05)
  model_psa2[i, "u_death"]  <- model_pars$u_death
  model_psa2[i, "c_pre_trt"]<- rlnorm(1, meanlog = log(model_pars$c_pre_trt), sdlog = 0.3)
  model_psa2[i, "c_pre_ref"]<- rlnorm(1, meanlog = log(model_pars$c_pre_ref), sdlog = 0.8)
  model_psa2[i, "c_post"]   <- model_pars$c_post
  model_psa2[i, "c_death"]   <- model_pars$c_death


  # Calculate the Markov traces
  pars_trt_i <- list(
    scale.os  = model_psa2[i, "OS_scale"],
    scale.pfs = model_psa2[i, "PFS_scale"],
    shape     = model_psa2[i, "shape"]
  )
  trace_trt_i <- get_partsurv(t_eval, pars_trt_i)

  pars_ref_i <- list(
    scale.os  = model_psa2[i, "OS_scale"] * (1 / model_psa2[i, "HR_OS"]),
    scale.pfs = model_psa2[i, "PFS_scale"]* (1 / model_psa2[i, "HR_PFS"]),
    shape     = model_psa2[i, "shape"]
  )
  trace_ref_i <- get_partsurv(t_eval, pars_ref_i)

  # Do the economic evaluation
  ee_trt_i <- get_ee(t_eval, 
                   trace_trt_i,
                   costs_per_cycle = model_psa2[i, c("c_pre_trt", "c_post", "c_death")],
                   utils           = model_psa2[i, c("u_pre_trt", "u_post", "u_death")],
                   dr              = model_settings$disc_rate,
                   cycle_length    = model_settings$cycle_length
  )
  
  ee_ref_i <- get_ee(t_eval, 
                     trace_ref_i,
                     costs_per_cycle = model_psa2[i, c("c_pre_ref", "c_post", "c_death")],
                     utils           = model_psa2[i, c("u_pre_ref", "u_post", "u_death")],
                     dr              = model_settings$disc_rate,
                     cycle_length    = model_settings$cycle_length
  )
  
  # Store results
  model_psa2[i, c("costs.trt", "QALYs.trt")] <- ee_trt_i
  model_psa2[i, c("costs.ref", "QALYs.ref")] <- ee_ref_i
  model_psa2[i, c("dCost", "dQALY")]         <- ee_trt_i - ee_ref_i
  model_psa2[i, c("ICER")]                   <- model_psa2[[i, "dCost"]] / model_psa2[[i, "dQALY"]]
}

ggplot(data = as.data.frame(model_psa2)) +
  geom_point(aes(dQALY, dCost), shape = 1) +
  geom_point(data = as.data.frame(t(delta)), 
             aes(delta.total.QALYs, delta.total.costs), col = "red") +
  geom_hline(yintercept = 0, lwd = 1.5) +
  geom_vline(xintercept = 0, lwd = 1.5) +
  geom_abline(intercept = 0, slope = 50000, linetype = "dashed") +
  scale_x_continuous(limits = c(-1, 5.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(-50000, 300000), 
                     breaks = seq(-50000, 300000, 50000), expand = c(0, 0)) +
  theme_bw()
```


## Summarize PSA outputs and (re)analyze the PSA outputs with the BCEA package
```{r, echo=TRUE}
# PSA outputs
e2 <- model_psa2[, c("QALYs.trt", "QALYs.ref")] # ref in bcea terminology is our trt!!
c2 <- model_psa2[, c("costs.trt", "costs.ref")] # caution to this !!
me2 <- apply(e2, MAR = 2, FUN = mean)
me2
mc2 <- apply(c2, MAR = 2, FUN = mean)
mc2

diff(mc2) / diff(me2)

# Compare with basecase analysis
ee_trt
ee_ref
delta
ICER


# re-analyze with bcea
res2 <- bcea(e2, c2, interventions = c("ref", "trt"), 
            Kmax = 100000, wtp = seq(0, 100000, 2000))

summary.bcea(res2, wtp = 50000)
```

```{r}
plot(res2, wtp = 50000)
```


## Calculate EVPPI
```{r}
par(mfrow = c(1,1))
# tte pars
voi2 <- evppi(c("HR_OS", "HR_PFS"), model_psa2, res2)
plot(voi2)

voi2 <- evppi(c("OS_scale", "PFS_scale"), model_psa2, res2)
plot(voi2)


# costs
voi2 <- evppi(c("c_pre_trt", "c_pre_ref"), model_psa2, res2)
plot(voi2)

# utils
voi2 <- evppi(c("u_pre_trt", "u_pre_ref"), model_psa2, res2)
plot(voi2)
```




# Session info
```{r}
getwd()
sessionInfo()
```

